---
title: "HW2"
author: "Haoran Zhang"
date: "2021/10/6"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(data.table)
library(dplyr)
library(dtplyr)
library(tidyverse)
library(ggplot2)
library(leaflet)
```


## Data Wrangling
```{r}
individual<-fread("chs_individual.csv")
regional<-fread("chs_regional.csv")
```

```{r}
total<-merge(x=individual,y=regional,by="townname",all.x = TRUE,all.y = FALSE)
```
Checking if there is any duplication.
```{r}
nrow(regional)
nrow(individual)
nrow(total)<nrow(regional)+nrow(individual)
```
Find those variables with NAs'.
```{r}
names(which(colSums(is.na(total)) > 0))
```
```{r}
class(total$height)="double"
```

We only need to compute the NA's of these numeric variables, while meaningless to input the NA's of dichotomous variables 
```{r}
total[, agepft := fcoalesce(agepft, mean(agepft, na.rm = TRUE)), by = .(male, hispanic)]
total[, height := fcoalesce(height, mean(height, na.rm = TRUE)), by = .(male, hispanic)]
total[, bmi := fcoalesce(bmi, mean(bmi, na.rm = TRUE)), by = .(male, hispanic)]
total[, fev := fcoalesce(fev, mean(fev, na.rm = TRUE)), by = .(male, hispanic)]
total[, fvc := fcoalesce(fvc, mean(fvc, na.rm = TRUE)), by = .(male, hispanic)]
total[, mmef := fcoalesce(mmef, mean(mmef, na.rm = TRUE)), by = .(male, hispanic)]
total[, no_24hr := fcoalesce(no_24hr, mean(no_24hr, na.rm = TRUE)), by = .(male, hispanic)]
total[, pm2_5_fr := fcoalesce(pm2_5_fr, mean(pm2_5_fr, na.rm = TRUE)), by = .(male, hispanic)]
```
## 2. Create a new categorical variable named “obesity_level”
```{r}
total[,obesity_level:=fifelse(bmi<14,"underweight",
                       fifelse(bmi<22,"normal",
                        fifelse(bmi<24,"overweight",
                               "obese")))]
table(total$obesity_level,useNA = "always")
table<-total[,.(maximum=max(bmi),minimum=min(bmi),total_no=.N), by=obesity_level]
knitr::kable(table)
```
## 3. Create another categorical variable named “smoke_gas_exposure” 
```{r}
smoke_gas<-total[,smoke_gas_exposure:=fifelse(smoke==0 & gasstove==0,"no_exposure",
                                   fifelse(smoke==0 & gasstove==1,"only_gas",
                                           fifelse(smoke==1 & gasstove==0, "only_smoke",
                                                   "both")))]
table(total$smoke_gas)
```
## 4. Create four summary tables showing the average 
```{r}
table_townname <- total[, .(
  fev_mean    = mean(fev,na.rm = TRUE),
  fev_sd     = sd(fev,na.rm = TRUE),
  asthma_mean = mean(asthma,na.rm = TRUE),
  asthma_sd = sd(asthma,na.rm = TRUE)
), by = townname]
knitr::kable(table_townname)
```

```{r}
table_sex <- total[, .(
  fev_mean    = mean(fev,na.rm = TRUE),
  fev_sd     = sd(fev,na.rm = TRUE),
  asthma_mean = mean(asthma,na.rm = TRUE),
  asthma_sd = sd(asthma,na.rm = TRUE)
), by = male]
knitr::kable(table_sex)
```

```{r}
table_obesity <- total[, .(
  fev_mean    = mean(fev,na.rm = TRUE),
  fev_sd     = sd(fev,na.rm = TRUE),
  asthma_mean = mean(asthma,na.rm = TRUE),
  asthma_sd = sd(asthma,na.rm = TRUE)
), by = obesity_level]
knitr::kable(table_obesity)
```

```{r}
table_sme <- total[, .(
  fev_mean    = mean(fev,na.rm = TRUE),
  fev_sd     = sd(fev,na.rm = TRUE),
  asthma_mean = mean(asthma,na.rm = TRUE),
  asthma_sd = sd(asthma,na.rm = TRUE)
), by = smoke_gas_exposure]
knitr::kable(table_sme)
```
## Looking at the Data (EDA)

## 1.Facet plot showing scatterplots with regression lines of BMI vs FEV by “townname”.
```{r}
total[!is.na(bmi) & !is.na(fev)] %>% 
  ggplot() + 
  geom_point(mapping = aes(x = bmi, y = fev, color=townname)) + 
  geom_smooth(method = lm, mapping = aes(x = bmi, y = fev))+
  facet_wrap(~ townname, nrow = 3)
```
fev gets larger as bmi increases at every location.


## 2. Stacked histograms of FEV by BMI category and FEV by smoke/gas exposure. Use different color schemes than the ggplot default.
```{r}
total[!is.na(fev) & !is.na(obesity_level)] %>%
  ggplot() + 
  geom_histogram(mapping = aes(x = fev, fill = obesity_level))+
  scale_fill_viridis_d(option = "plasma")
```
The distribution is basically normal, while underweight population has relatively lower fev, normal obesity population has medium fev and overweighted population has relatively larger fev. In conclusion, people with higher obesity level tend to have larger fev.
```{r}
total[!is.na(fev) & !is.na(smoke_gas_exposure)] %>%
  ggplot() + 
  geom_histogram(mapping = aes(x = fev, fill = smoke_gas_exposure))+
  scale_fill_viridis_d(option="magma")
```
The distribution is normal, but no other relation can be interpreted between smoke gas exposure and fev.

## 3. Barchart of BMI by smoke/gas exposure.
```{r}
total[!is.na(bmi)&!is.na(smoke_gas_exposure)] %>%
  ggplot() + 
  geom_bar(mapping = aes(x = obesity_level, colour = smoke_gas_exposure, fill=smoke_gas_exposure))+
  scale_fill_viridis_d(option="mako")
```
People with only gas exposure have the largest proportion, while population with normal obesity level have a relatively large proportion of no exposure compare with the other three obesity levels.
## 4. Statistical summary graphs of FEV by BMI and FEV by smoke/gas exposure category.
```{r}
total[!is.na(fev)] %>%
  ggplot() + 
    stat_summary(mapping = aes(x = obesity_level, y = fev),
    fun.min = min,
    fun.max = max,
    fun = median)+
  labs(title = "statistical summary of FEV by BMI category")
```
Again, fev gets larger as obesity level increases.

```{r}
total[!is.na(fev)& !is.na(smoke_gas_exposure)] %>%
  ggplot() + 
    stat_summary(mapping = aes(x = smoke_gas_exposure, y = fev),
    fun.min = min,
    fun.max = max,
    fun = median)+
  labs(title = "statistical summary of FEV by smoke gas exposure")
```
Not quite an association between smoke gas exposure and fev.

## 5. A leaflet map showing the concentrations of PM2.5 mass in each of the CHS communities.
```{r}
pal <- colorFactor(c("red","blue"), domain = total$pm25_mass)
leaflet() %>% 
  addProviderTiles('CartoDB.Positron') %>% 
  addCircles(
    data = total,
    lat = ~lat, lng = ~lon, 
    opacity = 1, fillOpacity = 1, radius = 400, color = ~pal(total$pm25_mass)
    ) %>%
  addLegend('bottomleft', pal=pal, values=total$pm25_mass,
          title='PM2.5 Concentration', opacity=1)
```
Seems that PM2.5 concentration is high around Jurupa Valley. 
## 6. Choose a visualization to examine whether PM2.5 mass is associated with FEV.
```{r}
total[!is.na(fev)&!is.na(pm25_mass)] %>%
  ggplot(mapping = aes(x = pm25_mass, y = fev, color = townname, linetype = townname)) +
  geom_point() + 
  geom_smooth(method = lm, se = FALSE, col = "black")
```
Seems that there is no association between pm2.5 mass and fev.


